### 2025-09-23 paper/sec/introduction.tex

AI-assisted authoring of transparent documents thus support turning static text into interactable, data-driven
content able to expose the evidential basis of scholarly claims, for which we envisage two use cases. First,
when \textbf{authoring} content for an online article, a journalist or scientific publisher may wish to
provide text which is linked to the underlying data so that the evidence base for the textual claims can be
explored directly from the article. Second, when \textbf{reading} a document reporting on findings derived
from open data (perhaps a scientific paper or climate report), the reader may want to retroactively interpret
parts of the text as queries over the available data and gradually ``rationally reconstruct'' the relationship
between claims in the paper and the evidence base. This might be just to aid their own comprehension, or part
of a formal peer review process.

>>>>

AI-assisted authoring of transparent documents thus support turning static text into interactable, data-driven
content able to expose the evidential basis of scholarly claims. We envisage two use cases. First, when
\textbf{authoring} content for an online article, a journalist or scientific publisher may wish to provide
text which is linked to the underlying data so that the evidence base for the textual claims can be explored
directly from the article. Second, when \textbf{reading} a document reporting on findings derived from open
data (perhaps a scientific paper or climate report), the reader may want to retroactively interpret parts of
the text as queries over the available data and gradually ``rationally reconstruct'' the relationship between
claims in the paper and the evidence base. This might be just to aid their own comprehension, or part of a
formal peer review process.
